---
title: "ODEs"
author: "spsaaibi"
date: "March 24, 2015"
output:
  html_document:
    highlight: espresso
    theme: readable
---

Differential equations can be separated into stochastic versus deterministic DEs. Problems can be split into initial value problems versus boundary value problems. One also distinguishes ordinary differential equations (ODEs) from partial differential equations (PDEs), differential algebraic equations (DAE) and delay differential equations (DDE). All these types of DEs can be solved in R. DE problems can be classified to be either stiff or nonstiff; the former type of problems are much more difficult to solve.



```{r, echo=FALSE}
library(deSolve)
library(ggplot2)
```

We will aim to solve the following equation $\frac{dy}{dx} = y^{\prime} = f(x,y)$

```{r, echo = FALSE}
r <- 1 ; K <- 10
 yini <- c(y = 2)
 derivs <- function(t, y, parms)
    list(r * y * (1-y/K))
 times <- seq(from = 0, to = 20, by = 0.2)
 out   <- ode(y = yini, times = times, func = derivs,
             parms = NULL)
 head(out, n = 3)
```


```{r}
yini <- c(y = 12)
 out2 <- ode(y = yini, times = times, func = derivs,
             parms = NULL)
 plot(out, out2, main = "logistic growth", lwd = 2)
```



## ODEs basics: Chaos in the Atmosphere

The Lorenz equations (Lorenz, 1963) were the first chaotic dynamic system to be described. They consist of three differential equations that were assumed to represent idealized behavior of the earth’s atmosphere. We use this model to demonstrate how to implement and solve differential equations in R. The Lorenz model describes the dynamics of three state variables, X, Y and Z. The model equations are:

